---
title: "Supporting Information"
subtitle: >
  `Herbivore-induced volatile signalling is conserved across \ locally adapted populations of _Arabidopsis thaliana_`{=typst}
authors:
  Rodrigo R. Granjel`#super[1]`{=typst}*,
  Lucía Martín-Cacheda`#super[2,3]`{=typst}
  Gregory Röder`#super[4]`{=typst},
  Iago Izquierdo-Ferreiro`#super[5]`{=typst},
  Andrea Martín-Díaz`#super[6]`{=typst},
  F. Xavier Picó`#super[6]`{=typst}

format: 
  typst: 
    mainfont: Roboto
    header-includes: |
      #show heading: set block(below: 2em)

execute: 
  echo: false
  message: false
  warning: false

bibliography: supp-info/grateful-refs.bib
csl: supp-info/new-phytologist.csl
---

`#v(0.75cm)`{=typst}

`#super[1]Basque Centre for Climate Change (BC3), 48940, Leioa, Spain; #super[2]Department of Ecology, Environment and Plant Sciences, Stockholm University, SE, 10691, Stockholm, Sweden; #super[3]Misión Biológica de Galicia (MBG-CSIC), Apartado de correos 28, Pontevedra, Galicia 36080, Spain; #super[4]Institute of Biology, University of Neuchâtel, Rue Emile-Argand 11, Neuchâtel 2000, Switzerland; #super[5]Centro Oceanográfico de Vigo, Instituto Español de Oceanografía (IEO, CSIC), Subida a Radio Faro 50, 36390 Vigo, Spain; #super[6]Departamento de Ecología y Evolución, Estación Biológica de Doñana (EBD), Consejo Superior de Investigaciones Científicas (CSIC), 41092 Sevilla, Spain \ \`{=typst}

*Corresponding email: [granjel@gmail.com](mailto:granjel@gmail.com)

<!-- `#v(1.5cm)`{=typst}

**This document contains:**

- Tables SX
- Figures SX
- Software citations -->

```{r load-all-data}
# load all necessary scripts and objects for the supporting information
results_permanova <- read.csv("tables/permanova/permanova-vocs-all.csv")
source("code/12-soil-samples.R")
```

{{< pagebreak >}}

# Supporting tables

**Table S1.** Results of the PERMANOVA analysis testing for differences in volatile organic compound (VOC) profiles between populations (Bon and Cai), treatments (control and herbivore-induced), and their interaction. The table includes degrees of freedom (Df), sum of squares (SS), R², F value, and p value for each factor and their interaction.

```{r table-permanova}
results_permanova <- results_permanova %>%
  mutate(
    # rename factors for better table formatting
    factor = case_when(
      factor == "population" ~ "Population",
      factor == "treatment" ~ "Treatment",
      factor == "treatment:population" ~ "Population x Treatment",
      TRUE ~ factor
    )
  ) %>%
  # reorder rows explicitly
  slice(match(
    c("Population", "Treatment", "Population x Treatment", "Residual", "Total"),
    factor
  )) %>%
  mutate(
    # round values
    Pr..F. = {
      p <- round(Pr..F., 3)
      sig <- case_when(
        p < 0.001 ~ "***",
        p < 0.01 ~ "**",
        p < 0.05 ~ "*",
        TRUE ~ ""
      )
      ifelse(
        is.na(p),
        "",
        paste0(
          format(p, nsmall = 3, trim = TRUE),
          ifelse(sig == "", "", paste0(" ", sig))
        )
      )
    },
    F = round(F, 2),
    R2 = round(R2, 3),
    SumOfSqs = round(SumOfSqs, 3)
  ) %>%
  mutate(across(everything(), ~ ifelse(is.na(.), "", .)))

table_permanova <- kable(
  results_permanova,
  col.names = c(
    "",
    "Df",
    "SS",
    "R²",
    "F",
    "p"
  ),
  digits = c(0, 0, 2, 3, 2, 4),
  align = "c",
  booktabs = TRUE
)

table_permanova
```

{{< pagebreak >}}

# Supporting figures

```{r plot-vocs-experiment}
#| fig-align: center
#| fig-width: 8
#| fig-height: 9.25
# print plot
plot_vocs_experiment
```

**Figure S1.** Volatile organic compounds (VOCs) from _Arabidopsis thaliana_ plants in the experiment, separated by population and treatment. Dots represent mean emissions and error bars represent the 5th and 95th percentiles. Different colors indicate different VOC types.

{{< pagebreak >}}

```{r plot-soil-blanks}
#| fig-align: center
#| fig-width: 6
#| fig-height: 6
# print plot
plot_soil_blanks
```

**Figure S2.** Volatile organic compounds (VOCs) from blank samples (no plants, only seedling wells with soil), collected to account for background emissions not produced by plants. Dots represent mean emissions and error bars represent the 5th and 95th percentiles. Different colors indicate different VOC types.

{{< pagebreak >}}

```{r plot-lob-perc}
#| fig-align: center
#| fig-width: 9
#| fig-height: 7.5
# print plot
plot_lob_perc
```

**Figure S3.** Percentage of VOC measurements above the limit of blanks for each compound across maternal lines, for each population and treatment, with the different VOC types indicated by color.

{{< pagebreak >}}

# Software citations

```{r}
#| echo: false
grateful::cite_packages(
  output = "paragraph",
  pkgs = "Session",
  out.dir = "supp-info",
  citation.style = "new-phytologist"
)
```

{{< pagebreak >}}

