---
title: "Supporting Information"
subtitle: >
  `Herbivore-induced volatile signalling is conserved across \ locally adapted populations of _Arabidopsis thaliana_`{=typst}
authors:
  Rodrigo R. Granjel`#super[1]`{=typst}*,
  Lucía Martín-Cacheda`#super[2,3]`{=typst}
  Gregory Röder`#super[4]`{=typst},
  Iago Izquierdo-Ferreiro`#super[5]`{=typst},
  Andrea Martín-Díaz`#super[6]`{=typst},
  F. Xavier Picó`#super[6]`{=typst}

format: 
  typst: 
    mainfont: Roboto
    header-includes: |
      #show heading: set block(below: 2em)

crossref:
  fig-title: Figure S
  fig-prefix: Fig. S

execute: 
  echo: false
  message: false
  warning: false

bibliography: supp-info/grateful-refs.bib
csl: supp-info/new-phytologist.csl
---

`#v(0.75cm)`{=typst}

`#super[1]Basque Centre for Climate Change (BC3), 48940, Leioa, Spain; #super[2]Department of Ecology, Environment and Plant Sciences, Stockholm University, SE, 10691, Stockholm, Sweden; #super[3]Misión Biológica de Galicia (MBG-CSIC), Apartado de correos 28, Pontevedra, Galicia 36080, Spain; #super[4]Institute of Biology, University of Neuchâtel, Rue Emile-Argand 11, Neuchâtel 2000, Switzerland; #super[5]Centro Oceanográfico de Vigo, Instituto Español de Oceanografía (IEO, CSIC), Subida a Radio Faro 50, 36390 Vigo, Spain; #super[6]Departamento de Ecología y Evolución, Estación Biológica de Doñana (EBD), Consejo Superior de Investigaciones Científicas (CSIC), 41092 Sevilla, Spain \ \`{=typst}

*Corresponding email: [granjel@gmail.com](mailto:granjel@gmail.com)

<!-- `#v(1.5cm)`{=typst}

**This document contains:**

- Tables SX
- Figures SX
- Software citations -->

```{r load-all-data}
# load all necessary scripts
source("code/16-soil-samples.R")
```

{{< pagebreak >}}

# Supporting tables
...

{{< pagebreak >}}

# Supporting figures

```{r plot-soil-blanks}
#| label: fig-soil-blanks
#| fig-cap: "Volatile organic compounds (VOCs) from blank samples (no plants, only seedling wells with soil), collected to account for background emissions not produced by plants."
#| fig-alt: "A figure showing the mean abundance of VOCs in soil samples, with error bars representing the 5th and 95th percentiles. The x-axis represents the abundance in ng h^-1, and the y-axis lists the VOC compounds. Different colors indicate different VOC types."
#| fig-align: center
#| fig-width: 6
#| fig-height: 6
# print plot
plot_soil_blanks
```

{{< pagebreak >}}

```{r plot-lob-perc}
#| label: fig-lob-perc
#| fig-cap: "Percentage of VOC measurements above the limit of blanks for each compound across maternal lines, for each population and treatment, with the different VOC types indicated by color."
#| fig-alt: "A figure showing the percentage of VOC measurements above the limit of blanks for each compound across maternal lines, for each population and treatment. The x-axis represents the percentage of measurements above the limit of blanks, and the y-axis lists the VOC compounds. Different colors indicate different VOC types."
#| fig-align: center
#| fig-width: 9
#| fig-height: 7.5
# print plot
plot_lob_perc
```

{{< pagebreak >}}

# Software citations

```{r}
#| echo: false
grateful::cite_packages(
  output = "paragraph",
  pkgs = "Session",
  out.dir = "supp-info",
  citation.style = "new-phytologist"
)
```

{{< pagebreak >}}

